---
title: "C3T3"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
require(pacman)

pacman:: p_load(pacman, dplyr, GGally, ggplot2, ggrepel, patchwork, gifski, ggforce, ggthemes, maps, sf, concaveman, remotes, readxl, ggthemes, ggvis, httr, plotly, rmarkdown, extrafont, shiny, isoband, stringr, rio, tidyr, labeling, caret, jquerylib, farver, corrgram, caTools, cowplot, randomForest, RMariaDB, lubridate, zoo, scales, ggfortify, forecast, doParallel, e1071, explore, dlookr, corrplot)
```

###Objective:  
  
They have asked our team to analyze historical sales data and then make sales volume predictions for a list of new product types  
  
- Predicting sales of four different product types: PC, Laptops, Netbooks and Smartphones.  
- Assessing the impact services reviews and customer reviews have on sales of different product types.  
###Importing Data
```{r}
df1 <- import("existingproductattributes2017.csv")
df2 <- import("newproductattributes2017.csv")

#str(df1)
#names(df1)

df1 <- select(df1, -c(ProductNum, BestSellersRank, ProductWidth, ProductHeight, ProductDepth,ShippingWeight,ProfitMargin))

df2 <- select(df2, -c(ProductNum, BestSellersRank, ProductWidth, ProductHeight, ProductDepth,ShippingWeight,ProfitMargin))

#names(df1)
```

###Correlation Plot

```{r}
Dummy <- dummyVars(" ~ .", data = df1)
df11 <- data.frame(predict(Dummy, newdata = df1))

Dummy2 <- dummyVars(" ~ .", data = df2)
df22 <- data.frame(predict(Dummy, newdata = df2))
#is.na(df11)
#explore(df11)
#df11
corrplot(cor(df11), method = "square", tl.cex=0.5)
```
###Building Models
```{r, results="hide"}
set.seed(107)
inTrain <- createDataPartition(y = df11$Volume, p = .75, list = FALSE)
training <- df11[ inTrain,]
testing <- df11[-inTrain,]

ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3, classProbs = FALSE)

RF <- train(Volume~., data = training, method = "rf", tuneLength = 2, trControl=ctrl)
GBM = train(Volume ~., data=training, method="gbm", trControl=ctrl)
SVM <- train(Volume~., data = training, method = "svmLinear", trControl=ctrl, tuneLength = 5)
```


```{r}
importanceRF = varImp(RF, scale=TRUE)
importanceRF
```
###RMSE of 1st Models

```{r}
RF
GBM
SVM
```

###2nd Try at Modelling

What if i get rid of the productype and price which dont have much impact/importance on the modelling?
Second round of modelling:
```{r, results="hide"}
RF2 <- train(Volume~ x5StarReviews + x4StarReviews  + x3StarReviews  + x2StarReviews + x1StarReviews + PositiveServiceReview + NegativeServiceReview + Recommendproduct + Price, data = training, method = "rf", tuneLength = 2, trControl=ctrl)

GBM2 = train(Volume ~x5StarReviews + x4StarReviews  + x3StarReviews  + x2StarReviews + x1StarReviews + PositiveServiceReview + NegativeServiceReview + Recommendproduct + Price, data=training, method="gbm", trControl=ctrl)

SVM2 <- train(Volume~x5StarReviews + x4StarReviews  + x3StarReviews  + x2StarReviews + x1StarReviews + PositiveServiceReview + NegativeServiceReview + Recommendproduct + Price, data = training, method = "svmLinear", trControl=ctrl, tuneLength = 5)
```


```{r}
importanceRF2 = varImp(RF2, scale=TRUE)
importanceRF2
```
###RMSE of 2nd Models

```{r}
RF2
GBM2
SVM2
```

Since it does not make much of a difference, I will keep the original models and make predictions to assess the quality of each model.

###Predicions

```{r}
RFpred <- predict(RF, newdata = testing)
GBMpred <- predict(GBM, newdata = testing)
SVMpred <- predict(SVM, newdata = testing)

RFpred
GBMpred
SVMpred

```
###Applying Predictions

```{r}
RFpred2 <- predict(RF, newdata = df22)
```

```{r}
output <- df2 
output$predictions <- RFpred2

head(output)

#write.csv(output, file="C3T3 Predictions.csv", row.names = TRUE)
```

