---
title: "C5T3"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
require(pacman)

pacman:: p_load(pacman, dplyr, GGally, ggplot2, ggrepel, patchwork, gifski, ggforce, ggthemes, maps, sf, concaveman, remotes, readxl, ggthemes, ggvis, httr, plotly, rmarkdown, extrafont, shiny, isoband, stringr, rio, tidyr, labeling, caret, jquerylib, farver, corrgram, caTools, cowplot, randomForest, RMariaDB, lubridate, zoo, scales, ggfortify, forecast, doParallel,e1071,C50,kknn)
```
## Importing and Feature Selection

First step is to import the data and select the features that only apply to each phone.

###Original Data
```{r}
df_iphone <- import("iphone_smallmatrix_labeled_8d.csv")
df_samsung <- import("galaxy_smallmatrix_labeled_9d.csv")
```

### Histogram for EDA

```{r}
#plot_ly(df_iphone, x= ~df_iphone$iphonesentiment, type='histogram')
#plot_ly(df_samsung, x= ~df_samsung$galaxysentiment, type='histogram')
```
Where:  
0: Sentiment Unclear  
1: very negative  
2: somewhat negative  
3: neutral  
4: somewhat positive  
5: very positive  

###Domain Expertise Feature Selection
```{r}
df_iphoneDE <- select(df_iphone, c(ios, iphonecampos, iphonecamneg, iphonecamunc, iphonedispos, iphonedisneg, iphonedisunc, iphoneperpos,iphoneperneg,iphoneperunc,iosperpos,iosperneg,iosperunc,iphonesentiment))

df_samsungDE <- select(df_samsung, c(googleandroid, samsungcampos, samsungcamneg, samsungcamunc, samsungdispos, samsungdisneg, samsungdisunc, samsungperpos,samsungperneg,samsungperunc,googleperpos,googleperneg,googleperunc,galaxysentiment))
```

###Correlation Feature Selection
```{r}
Corrmatrix = cor(df_iphone)
Corrmatrix2 = cor(df_samsung)

#findCorrelation(Corrmatrix, cutoff = .8, verbose = TRUE, names = TRUE)
#findCorrelation(Corrmatrix2, cutoff = .8, verbose = TRUE, names = TRUE)

df_iphoneCOR <- select(df_iphone, c(samsungdisneg, samsungperneg, samsungdispos, htcdisneg,googleperneg, googleperpos, samsungdisunc, samsungcamunc, htcperpos,nokiacamunc,  nokiadisneg, nokiadispos, nokiaperunc, nokiacampos, nokiadisunc,nokiaperneg,nokiacamneg,iphonedisneg, 
iphonedispos,sonydispos, iosperunc, iosperneg, ios, htcphone,iphonesentiment))

df_samsungCOR <- select(df_samsung,c(samsungdisneg,samsungperneg,samsungdispos,htcdisneg,googleperneg,googleperpos,samsungdisunc,samsungcamunc,htcperpos,nokiacamunc,nokiadisneg,nokiadispos,nokiaperunc,nokiacampos,nokiadisunc,nokiaperneg,nokiacamneg,iphonedisneg,iphonedispos,sonyperpos,iosperunc,iosperneg,sonydisneg,ios,htcphone,galaxysentiment))

```

###NZV Feature Selection

Near zero variance feature selection

```{r}
nzv_iphone <- nearZeroVar(df_iphone, saveMetrics = FALSE)
nzv_samsung <- nearZeroVar(df_samsung, saveMetrics = FALSE)

df_iphoneNZV <- df_iphone[,-nzv_iphone]
df_samsungNZV <- df_samsung[,-nzv_samsung]

#str(df_iphoneNZV)
#str(df_samsungNZV)

```

### RFE Feature Selection 

Let's sample the data before using RFE
```{r}
set.seed(123)
iphoneSample <- df_iphone[sample(1:nrow(df_iphone), 1000, replace=FALSE),]
samsungSample <- df_samsung[sample(1:nrow(df_samsung), 1000, replace=FALSE),]
```

Set up rfeControl with randomforest, repeated cross validation and no updates
```{r}
ctrl <- rfeControl(functions = rfFuncs, method = "repeatedcv", repeats = 5, verbose = FALSE)
```


```{r}

#cl <- makeCluster(7)
#registerDoParallel(cl)

# Use rfe and omit the response variable (attribute 59 iphonesentiment) 
rfeResults_iphone <- rfe(iphoneSample[,1:58], iphoneSample$iphonesentiment, sizes=(1:58), rfeControl=ctrl)
rfeResults_samsung <- rfe(samsungSample[,1:58], samsungSample$galaxysentiment, sizes=(1:58), rfeControl=ctrl)
```


```{r}
# Get results
rfeResults_iphone
rfeResults_samsung
```

```{r}
#plot(rfeResults_iphone, type=c("g", "o"))
#plot(rfeResults_samsung, type=c("g", "o"))
```

```{r}
# create new data set with rfe recommended features
df_iphoneRFE <- df_iphone[,predictors(rfeResults_iphone)]
df_samsungRFE <- df_samsung[,predictors(rfeResults_samsung)]

# add the dependent variable to iphoneRFE
df_iphoneRFE$iphonesentiment <- df_iphone$iphonesentiment
df_samsungRFE$galaxysentiment <- df_samsung$galaxysentiment


```

## All Data Sets

df_iphone  Original Data  
df_samsung  

df_iphoneDE  Domain Expertise  
df_samsungDE  

df_iphoneCOR  Correlation  
df_samsungCOR  

df_iphoneNZV  Near zero variance feature selection  
df_samsungNZV  

df_iphoneRFE  Recursive Feature Elimination  
df_samsungRFE  

```{r}
df_iphone$iphonesentiment <- factor(df_iphone$iphonesentiment)
df_samsung$galaxysentiment <- factor(df_samsung$galaxysentiment)

df_iphoneDE$iphonesentiment <- factor(df_iphoneDE$iphonesentiment)
df_samsungDE$galaxysentiment <- factor(df_samsungDE$galaxysentiment)

df_iphoneCOR$iphonesentiment <- factor(df_iphoneCOR$iphonesentiment)
df_samsungCOR$galaxysentiment <- factor(df_samsungCOR$galaxysentiment)

df_iphoneNZV$iphonesentiment <- factor(df_iphoneNZV$iphonesentiment)
df_samsungNZV$galaxysentiment <- factor(df_samsungNZV$galaxysentiment)

df_iphoneRFE$iphonesentiment <- factor(df_iphoneRFE$iphonesentiment)
df_samsungRFE$galaxysentiment <- factor(df_samsungRFE$galaxysentiment)
```


## Model Development

Models with non-feature selected datasets and then feature selected data sets.
C5.0, Random Forest, SVM, kknn

```{r}
set.seed(123)

#iPhone

inTrain <- createDataPartition(df_iphone$iphonesentiment, p=.70, list = FALSE)
training_iphone <- df_iphone[ inTrain,]
testing_iphone  <- df_iphone[-inTrain,]

inTrain <- createDataPartition(df_iphoneDE$iphonesentiment, p=.70, list = FALSE)
training_iphoneDE <- df_iphoneDE[ inTrain,]
testing_iphoneDE  <- df_iphoneDE[-inTrain,]

inTrain <- createDataPartition(df_iphoneCOR$iphonesentiment, p=.70, list = FALSE)
training_iphoneCOR <- df_iphoneCOR[ inTrain,]
testing_iphoneCOR  <- df_iphoneCOR[-inTrain,]

inTrain <- createDataPartition(df_iphoneNZV$iphonesentiment, p=.70, list = FALSE)
training_iphoneNZV <- df_iphoneNZV[ inTrain,]
testing_iphoneNZV  <- df_iphoneNZV[-inTrain,]

inTrain <- createDataPartition(df_iphoneRFE$iphonesentiment, p=.70, list = FALSE)
training_iphoneRFE <- df_iphoneRFE[ inTrain,]
testing_iphoneRFE  <- df_iphoneRFE[-inTrain,]

#Samsung

inTrain <- createDataPartition(df_samsung$galaxysentiment, p=.70, list = FALSE)
training_samsung <- df_samsung[ inTrain,]
testing_samsung  <- df_samsung[-inTrain,]

inTrain <- createDataPartition(df_samsungDE$galaxysentiment, p=.70, list = FALSE)
training_samsungDE <- df_samsungDE[ inTrain,]
testing_samsungDE  <- df_samsungDE[-inTrain,]

inTrain <- createDataPartition(df_samsungCOR$galaxysentiment, p=.70, list = FALSE)
training_samsungCOR <- df_samsungCOR[ inTrain,]
testing_samsungCOR  <- df_samsungCOR[-inTrain,]

inTrain <- createDataPartition(df_samsungNZV$galaxysentiment, p=.70, list = FALSE)
training_samsungNZV <- df_samsungNZV[ inTrain,]
testing_samsungNZV  <- df_samsungNZV[-inTrain,]

inTrain <- createDataPartition(df_samsungRFE$galaxysentiment, p=.70, list = FALSE)
training_samsungRFE <- df_samsungRFE[ inTrain,]
testing_samsungRFE  <- df_samsungRFE[-inTrain,]
```

####Original Data Models
```{r}

#cl <- makeCluster(7)
#registerDoParallel(cl)

ctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 3)


RF_iphone_Original <- train(iphonesentiment~., data = training_iphone, method = "rf",trControl=ctrl, tuneLength = 1)
RF_samsung_Original <- train(galaxysentiment~., data = training_samsung, method = "rf",trControl=ctrl, tuneLength = 1)

SVM_iphone_Original <- train(iphonesentiment~., data = training_iphone, method = "svmLinear",trControl=ctrl, tuneLength = 3)
SVM_samsung_Original <- train(galaxysentiment~., data = training_samsung, method = "svmLinear",trControl=ctrl, tuneLength = 3)

C50_iphone_Original <- train(iphonesentiment~., data = training_iphone,method="C5.0",trControl=ctrl, tuneLength = 3)
C50_samsung_Original <- train(galaxysentiment~., data = training_samsung,method="C5.0",trControl=ctrl, tuneLength = 3)

kknn_iphone_Original<- train(iphonesentiment~., data = training_iphone, method = "kknn", trControl = ctrl, tuneLength = 3)
kknn_samsung_Original <- train(galaxysentiment~., data = training_samsung, method = "kknn", trControl = ctrl, tuneLength = 3)
```

###Predictions
```{r}
RF_pred_iphone <- predict(RF_iphone_Original, newdata = testing_iphone)
C50_pred_iphone <- predict(C50_iphone_Original, newdata = testing_iphone)
SVM_pred_iphone <- predict(SVM_iphone_Original, newdata = testing_iphone)
KKNN_pred_iphone <- predict(kknn_iphone_Original, newdata = testing_iphone)

RF_pred_samsung <- predict(RF_samsung_Original, newdata = testing_samsung)
C50_pred_samsung <- predict(C50_samsung_Original, newdata = testing_samsung)
SVM_pred_samsung <- predict(SVM_samsung_Original, newdata = testing_samsung)
KKNN_pred_samsung <- predict(kknn_samsung_Original, newdata = testing_samsung)
```

```{r}
#Building a confusion matrix and using the predict function yields a similar accuracy and kappa
#result than the resampling function. Since resampling summarises these metrics in a simpler
#manner, i will use resampling but will leave this matrix as an example. 

cmRF_iphone <- confusionMatrix(RF_pred_iphone, testing_iphone$iphonesentiment)
cmRF_samsung <- confusionMatrix(RF_pred_samsung, testing_samsung$galaxysentiment)

cmRF_iphone
cmRF_samsung
```

#### Original Data Results
```{r}
ModelData_samsung <- resamples(list(RF_iphone = RF_iphone_Original, KKNN_iphone = kknn_iphone_Original, C50_iphone = C50_iphone_Original, SVM_iphone = SVM_iphone_Original, RF_Samsung = RF_samsung_Original, KKNN_Samsung = kknn_samsung_Original, C50_Samsung = C50_samsung_Original, SVM_Samsung = SVM_samsung_Original))

summary(ModelData_samsung)
```

###Feture Selected Data Set

```{r}
#RFE Feature Selection

RF_iphone_RFE <- train(iphonesentiment~., data = training_iphoneRFE, method = "rf",trControl=ctrl, tuneLength = 1)
RF_samsung_RFE <- train(galaxysentiment~., data = training_samsungRFE, method = "rf",trControl=ctrl, tuneLength = 1)

SVM_iphone_RFE <- train(iphonesentiment~., data = training_iphoneRFE, method = "svmLinear",trControl=ctrl, tuneLength = 3)
SVM_samsung_RFE <- train(galaxysentiment~., data = training_samsungRFE, method = "svmLinear",trControl=ctrl, tuneLength = 3)

C50_iphone_RFE <- train(iphonesentiment~., data = training_iphoneRFE,method="C5.0",trControl=ctrl, tuneLength = 3)
C50_samsung_RFE <- train(galaxysentiment~., data = training_samsungRFE,method="C5.0",trControl=ctrl, tuneLength = 3)

kknn_iphone_RFE<- train(iphonesentiment~., data = training_iphoneRFE, method = "kknn", trControl = ctrl, tuneLength = 3)
kknn_samsung_RFE <- train(galaxysentiment~., data = training_samsungRFE, method = "kknn", trControl = ctrl, tuneLength = 3)

#NZV Feature Selection

RF_iphone_NZV <- train(iphonesentiment~., data = training_iphoneNZV, method = "rf",trControl=ctrl, tuneLength = 1)
RF_samsung_NZV <- train(galaxysentiment~., data = training_samsungNZV, method = "rf",trControl=ctrl, tuneLength = 1)

SVM_iphone_NZV <- train(iphonesentiment~., data = training_iphoneNZV, method = "svmLinear",trControl=ctrl, tuneLength = 3)
SVM_samsung_NZV <- train(galaxysentiment~., data = training_samsungNZV, method = "svmLinear",trControl=ctrl, tuneLength = 3)

C50_iphone_NZV <- train(iphonesentiment~., data = training_iphoneNZV,method="C5.0",trControl=ctrl, tuneLength = 3)
C50_samsung_NZV <- train(galaxysentiment~., data = training_samsungNZV,method="C5.0",trControl=ctrl, tuneLength = 3)

kknn_iphone_NZV<- train(iphonesentiment~., data = training_iphoneNZV, method = "kknn", trControl = ctrl, tuneLength = 3)
kknn_samsung_NZV <- train(galaxysentiment~., data = training_samsungNZV, method = "kknn", trControl = ctrl, tuneLength = 3)
```

```{r}
#COR Feature Selection

RF_iphone_COR <- train(iphonesentiment~., data = training_iphoneCOR, method = "rf",trControl=ctrl, tuneLength = 1)
RF_samsung_COR <- train(galaxysentiment~., data = training_samsungCOR, method = "rf",trControl=ctrl, tuneLength = 1)

SVM_iphone_COR <- train(iphonesentiment~., data = training_iphoneCOR, method = "svmLinear",trControl=ctrl, tuneLength = 3)
SVM_samsung_COR <- train(galaxysentiment~., data = training_samsungCOR, method = "svmLinear",trControl=ctrl, tuneLength = 3)

C50_iphone_COR <- train(iphonesentiment~., data = training_iphoneCOR,method="C5.0",trControl=ctrl, tuneLength = 3)
C50_samsung_COR <- train(galaxysentiment~., data = training_samsungCOR,method="C5.0",trControl=ctrl, tuneLength = 3)

kknn_iphone_COR<- train(iphonesentiment~., data = training_iphoneCOR, method = "kknn", trControl = ctrl, tuneLength = 3)
kknn_samsung_COR <- train(galaxysentiment~., data = training_samsungCOR, method = "kknn", trControl = ctrl, tuneLength = 3)

#DE Feature Selection

RF_iphone_DE <- train(iphonesentiment~., data = training_iphoneDE, method = "rf",trControl=ctrl, tuneLength = 1)
RF_samsung_DE <- train(galaxysentiment~., data = training_samsungDE, method = "rf",trControl=ctrl, tuneLength = 1)

SVM_iphone_DE <- train(iphonesentiment~., data = training_iphoneDE, method = "svmLinear",trControl=ctrl, tuneLength = 3)
SVM_samsung_DE <- train(galaxysentiment~., data = training_samsungDE, method = "svmLinear",trControl=ctrl, tuneLength = 3)

C50_iphone_DE <- train(iphonesentiment~., data = training_iphoneDE,method="C5.0",trControl=ctrl, tuneLength = 3)
C50_samsung_DE <- train(galaxysentiment~., data = training_samsungDE,method="C5.0",trControl=ctrl, tuneLength = 3)

kknn_iphone_DE<- train(iphonesentiment~., data = training_iphoneDE, method = "kknn", trControl = ctrl, tuneLength = 3)
kknn_samsung_DE <- train(galaxysentiment~., data = training_samsungDE, method = "kknn", trControl = ctrl, tuneLength = 3)
```


###All Results
```{r}

ModelData_All_iPhone <- resamples(list(RF_iphone=RF_iphone_Original, 
                                    RF_iphone_DE=RF_iphone_DE, 
                                    RF_iphone_COR=RF_iphone_COR, 
                                    RF_iphone_NZV=RF_iphone_NZV, 
                                    RF_iphone_RFE=RF_iphone_RFE,
                                    
                                    C50_iphone=C50_iphone_Original, 
                                    C50_iphone_DE=C50_iphone_DE, 
                                    C50_iphone_COR=C50_iphone_COR, 
                                    C50_iphone_NZV=C50_iphone_NZV, 
                                    C50_iphone_RFE=C50_iphone_RFE,
                                    
                                    SVM_iphone=SVM_iphone_Original, 
                                    SVM_iphone_DE=SVM_iphone_DE, 
                                    SVM_iphone_COR=SVM_iphone_COR, 
                                    SVM_iphone_NZV=SVM_iphone_NZV, 
                                    SVM_iphone_RFE=SVM_iphone_RFE,
                                    
                                    kknn_iphone=kknn_iphone_Original, 
                                    kknn_iphone_DE=kknn_iphone_DE, 
                                    kknn_iphone_COR=kknn_iphone_COR, 
                                    kknn_iphone_NZV=kknn_iphone_NZV, 
                                    kknn_iphone_RFE=kknn_iphone_RFE
                                    ))

ModelData_All_Samsung <- resamples(list(RF_samsung=RF_samsung_Original, 
                                    RF_samsung_DE=RF_samsung_DE, 
                                    RF_samsung_COR=RF_samsung_COR, 
                                    RF_samsung_NZV=RF_samsung_NZV, 
                                    RF_samsung_RFE=RF_samsung_RFE,
                                    
                                    C50_samsung=C50_samsung_Original, 
                                    C50_samsung_DE=C50_samsung_DE, 
                                    C50_samsung_COR=C50_samsung_COR, 
                                    C50_samsung_NZV=C50_samsung_NZV, 
                                    C50_samsung_RFE=C50_samsung_RFE,
                                    
                                    SVM_samsung=SVM_samsung_Original, 
                                    SVM_samsung_DE=SVM_samsung_DE, 
                                    SVM_samsung_COR=SVM_samsung_COR, 
                                    SVM_samsung_NZV=SVM_samsung_NZV, 
                                    SVM_samsung_RFE=SVM_samsung_RFE,
                                    
                                    kknn_samsung=kknn_samsung_Original, 
                                    kknn_samsung_DE=kknn_samsung_DE, 
                                    kknn_samsung_COR=kknn_samsung_COR, 
                                    kknn_samsung_NZV=kknn_samsung_NZV, 
                                    kknn_samsung_RFE=kknn_samsung_RFE
                                    ))

```

```{r}
summary(ModelData_All_iPhone)
```

```{r}
summary(ModelData_All_Samsung)
```

##Large Matrices


```{r}
dfLM_iphone <- import("LargeMatrix_iphone.csv")
dfLM_samsung <- import("LargeMatrix_samsung.csv")

dfLM_iphoneRFE <- dfLM_iphone[,predictors(rfeResults_iphone)]
dfLM_iphoneRFE$iphonesentiment <- dfLM_iphone$iphonesentiment

#No feature selection for Samsung

LM_Pred_iphone <- predict(RF_iphone_RFE, newdata = dfLM_iphoneRFE)
LM_Pred_Samsung <- predict(C50_samsung_Original, newdata = dfLM_samsung)
```

```{r}
summary(LM_Pred_iphone)
```

```{r}
summary(LM_Pred_Samsung)
```

```{r}
summary(df_iphone$iphonesentiment)
```

```{r}
summary(df_samsung$galaxysentiment)
```

