---
title: "C3T4"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
require(pacman)

pacman:: p_load(pacman, dplyr, GGally, ggplot2, ggrepel, patchwork, gifski, ggforce, ggthemes, maps, sf, concaveman, remotes, readxl, ggthemes, ggvis, httr, plotly, rmarkdown, extrafont, shiny, isoband, stringr, rio, tidyr, labeling, caret, jquerylib, farver, corrgram, caTools, cowplot, randomForest, RMariaDB, lubridate, zoo, scales, ggfortify, forecast, doParallel, e1071, explore, dlookr, corrplot, arules,arulesViz)
```

```{r}
Transactions <- read.transactions("ElectronidexTransactions2017.csv", format = "basket", sep=",", skip = 0)
inspect(head(Transactions)) #cant use head, need inspect command for basket analysis
```
```{r}
itemFrequencyPlot( Transactions, topN=20, main = 'Absolute Item Frequency Plot', type="absolute", horiz=TRUE)
itemFrequencyPlot(Transactions, topN = 20, main = 'Relative Item Frequency Plot', type = "relative", ylab = "Item Frequency (Relative)")
```
#Finding Rules

apriori!

```{r}
rules <- apriori(Transactions, parameter = list(supp = 0.001, conf = 0.8, minlen=2, maxlen=5), control=list(verbose=FALSE))
```

```{r}
rules
inspect(rules[1:5])
```

Count = #baskets that satify rule  
Support = fraction of baskets that satisfy rule or fraction of baskets that have all items referenced in rule  
Confidence = (posterior) probability that cart has item on RHS given it has all items on LHS of rule  
coverage = fraction baskets that have all items in the LHS of rule.  
lift = how many times more likely that item in RHS appears in a cart when it contains all items in LHS  
RHS = right hand side  
LHS = left hand side  

```{r}
#Sorting by quality measure
inspect( sort(rules, by="lift", decreasing=TRUE)[1:5])
```


```{r}
inspect( sort(rules, by="confidence", decreasing=TRUE)[1:5])
```
```{r}
redundant<-is.redundant(rules)
redundant
```
```{r}
items <- itemFrequency(Transactions)
sorted<-sort(items,decreasing=TRUE)
Frequency<-sorted[1:20]
barchart(Frequency)
```


```{r}
#Visualize them

simplerules <- sort(rules, by="lift")[c(1:8)]
plot(simplerules, method="graph")
#plot(simplerules, method="graph", engine="htmlwidget")
```

```{r}

```

